Richness model

### 2024-07-30 julianna_richness_model.Rmd
Worked on richness model--made a first pass--maybe actually want to add an interaction?

```{r load data}
cleaned_rich <- read_csv(here("data/algal_metrics_quadrat.csv")) # generated using my (Julianna's) algal metric code
```

@@

```{r}
mRich <- glmer(,
               family = Poisson,
               data = cleaned_rich)
```

>>>>

```{r load data}
cleaned_rich <- read_csv(here("data/algal_metrics_quadrat.csv")) %>% 
  filter(year != 2020) # fake data
  # generated using my (Julianna's) algal metric code
```

@@

```{r make Poisson model}
mRich <- glmer(algal_richness ~ as.factor(year) + habitat + (1|site/transect),
               family = poisson,
               data = cleaned_rich)
```

# Examine output

```{r make prediction grid}
# create a grid to predict over
newdata <- expand_grid(year = unique(cleaned_rich$year),
                       habitat = unique(cleaned_rich$habitat),
                       site = unique(cleaned_rich$site),
                       transect = unique(cleaned_rich$transect))
```

```{r generate predictions}
newdata$Preds <- predict(mRich, newdata = newdata)
```

```{r plot predictions}
newdata %>% 
  # get transect means
  group_by(year, habitat, site) %>% 
  summarize(Pred_mean_richness = mean(Preds),
            SD_richness = sd(Preds)) %>% 
  ggplot(aes(x = year, y = Pred_mean_richness, group = habitat, color = habitat)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~site)
```

- Should year be a factor?
